The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2012-01-30 bmc_article.tex
Added a draft paragraph on the new builders

\title{The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, ...}

@@

>>>>

\title{The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching}

@@

Originally, the CDK was developed as a shared library between JChemPaint and Jmol. The former
used a MVC approach with an event-passing mechanism to update the view when the model was
changed. This can cause an cascade of change events being passed around. To address,
interfaces were introduced allowing multiple implementations of the core interfaces.
% TODO: check if the interfaces have previously been defined
The IChemObjectBuilders play an important role here, allowing implementations of the
interfaces to be instantiated without the need of explicitly referencing those implementations.

However, the CDK 1.0 and 1.2 implementations of the IChemObjectBuilder had one method for
each constructor, resulting in very large interface. Moreover, the API changed whenever
a new class was introduced, and existing methods changed when constructors were updated.
To simplify the API, the new IChemObjectBuilder collapsed all methods returning new
implementations into a single method, which takes as a first parameter the class of the
interface that is wished to be constructed. All further parameters are passes as
parameter to the class constructor.

For example, to construct a new atom from its element symbol, one would now write:

\begin{verbatim}
IChemObjectBuilder builder; // previously defined
IAtom atom = builder.newInstance(IAtom.class, "C");
\end{verbatim}

Increasingly, the CDK library is now written against the interfaces, and when new instanced
are needed, these builders are being used. This allows to run a certain CDK-based
application with a specific builder, summarized in Table~\ref{tab:builders}.

\begin{table}
\begin{tabular}{l|l}
builder & description \\
\hline
DefaultChemObjectBuilder & The original builder ... \\
DataDebugChemObjectBuilder & A builder that creates classes that send debug messages ... \\
SilentChemObjectBuilder & Classes that are created with this builder will not
    send around change events. \\
\end{tabular}
\end{table}
